Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench

### 2021-11-30 paper/paper.tex
Add text for sha3 overview
Add background on SAW and SHA-3

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{comment}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

@@

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\keccak{\textsc{Keccak}}

@@

\section{SHA-3 Overview}
\input{sha-3}

>>>>

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{comment}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

@@

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\keccak{\textsc{Keccak}}
\newcommand\shaThree{\texttt{SHA-3}}
\newcommand\nist{\texttt{NIST}}
\newcommand\fips{\texttt{FIPS 202}}
\newcommand\saw{\texttt{SAW}}
\newcommand\cryptol{Cryptol}
\newcommand\openssl{OpenSSL}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\algoref}[1]{Alg.~\ref{#1}}
\newcommand{\lineref}[1]{Line~\ref{#1}}

@@

\section{Background}\label{sec:background}
\cryptol\ and \saw\ are used to prove properties of \openssl's implementation of \shaThree\ relative to the \nist\ \fips\ publication \cite{fips202}. 
What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw, and (2) how the \shaThree\ algorithm constructs its digests.
Key elements relative to understanding the proof of each (\cryptol, \saw, and \shaThree) are defined in this section.

\subsection{Cryptol and SAW}\label{subsec:saw}
\input{saw}
\subsection{SHA-3 Overview}\label{subsec:sha3}
\input{sha-3}

### 2021-11-30 paper/proof.tex
Revise algorithm details and clarify what is proved
Add background on SAW and SHA-3

\begin{compactitem}
  \item Process 1600 bits at a time: $5 * 5 * 64$
  \item State vector is $5*5*64*$
  \item Incoming message padded out to whole number of 1600 bit bit-strings
  \item Processed in order by \texttt{KECCAK}
  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
  \item Keccak has 24 rounds for state vector. 
  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated 
  \item The max digest (hash) size is 512 bits in the standard
  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size 
\end{compactitem}

SAW Details
\begin{compactitem}
  \item Cryptol is a specification language specifically for defining crypto-primitives
  \item It has a compiler to turn Cryptol to SAW core terms
  \item It has an execution environment to run test vectors on specifications
  \item SAW is a tool to prove equivalence between SAW core terms
  \item It uses symbolic execution and must unroll every loop to a fixed length
  \item It can take as input C and Java code and compile those to core terms
  \item It can use overides to replace C and Java code in the term
\end{compactitem}

>>>>

\begin{compactitem}
  \item Process 1600 bits at a time: $5 * 5 * 64$
  \item State vector is $5*5*64*$
  \item Incoming message padded out to whole number of 1600 bit bit-strings
  \item Processed in order by \texttt{KECCAK}
  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
  \item Keccak has 24 rounds for state vector. 
  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated 
  \item The digest size is defined in pairs by the standard.
  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size
  \item The brute force proof is for a digest of 256-bits.
\end{compactitem}

SAW Details

### 2021-11-30 paper/sha-3.tex
Add background on SAW and SHA-3

\shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$-bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit hash (\lineref{line:squeeze}).
The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$-bits, which is initialized to 0 (\lineref{line:init}).

The state itself consists of $r$-bits that are written or read plus a capacity of $c = s - r$ that is untouched by input or output.
The \nist\ standard defines pairs of allowed hash sizes and rates as required in \algoref{alg:sha3}. The input $N$ is padded with $10^*1$ such that its final size is a multiple of $r$ bits, and it is partitioned into $r$-bit size blocks (\lineref{line:pad}) for the absorption phase.

>>>>

\shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$ bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit digest (\lineref{line:squeeze}).
The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$ bits, which is initialized to 0 (\lineref{line:init}).

The state itself consists of $r$ bits that are written or read plus a capacity of $c = s - r$ bits that are untouched by input or output.
The \nist\ standard defines pairs of allowed hash sizes and rates.
These are defined in the requires statement in \algoref{alg:sha3}. 
The input $N$ is padded such that its final size is a multiple of $r$ bits, and it is partitioned into $r$-bit size blocks (\lineref{line:pad}) for the absorption phase.
The padding is defined by the regular expression $10^*1$.
An extra block is added for the padding if needed since every $N$ must be a multiple of $r$-bit size blocks matching the $10^*1$ regular expression at the end of the last block.
